Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-01-04 RNARedPrint-Bioinfo2018.tex
Update on Overleaf.

\begin{figure*}[t]
{\centering\includegraphics[width=.8\textwidth]{Workflow}\\}
\caption{General outline of \ourprog{} for base pair-based energy models. From a set of target secondary structures (i), base-pairs are merged (ii) into a compatibility graph (iii) and transformed into a tree decomposition (iv). The tree is then used to compute the partition function, followed by a Boltzmann sampling of valid sequences (v). An adaptive scheme learns weights to achieve targeted energies and $\Gb\Cb$-content, leading to the production of suitable designs (vi).}
\label{fig:workflow}
\end{figure*}

@@

To illustrate our above strategy, we sampled $n=1\,000$ sequences
targeting ${\sf GC}_{target}=0.5$ and different Turner energies for
the three structure targets of an example
instance. Figure~\ref{fig:energydist}A illustrates how the Turner
energy distributions of the shown target structures can be accurately
shifted to prescribed target energies; for comparison, we plot the
respective energy distribution of uniformly and Boltzmann sampled
sequences. Figure~\ref{fig:energydist}B summarizes the targeting accuracy for the single structure energies over a larger set of instances
from the \texttt{Modena} benchmark. We emphasize that only the simple stacking energies
are directly targeted (at $\pm$10\% tolerance, expect for a few hard
instances). Due to our linear adjustment, we still achieve well
controlled energies in the ultimately relevant Turner energy model.
\begin{figure*}[t]
  \begin{center}
    {\sf \bfseries A}\includegraphicstop[width=0.57\textwidth]{Figs/energy_distribution}
    {\sf \bfseries B}\includegraphicstop[width=0.32\textwidth]{Figs/offset}
  \end{center}
  \caption{%
  	Targeting specific energies using multi-dimensional Boltzmann sampling. (A) Turner energy distributions for three target structures (right) while targeting $(-40,-40,-20)$ and $(-20,-20,-20)$ free energy vectors. For comparison, we show the distribution of uniform and Boltzmann samples; respectively associated with homogenous weights $1$ and $e^\beta$. (B) Accuracy of targeting over all target structures of the \Software{Modena} benchmark instances \texttt{RNAtabupath}, \texttt{3str} and \texttt{4str}. Shown is the relative deviation of targeted and achieved energies in the simple stacking model and the Turner model.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
}
  \label{fig:energydist}
\end{figure*}
\subsection{Generating high-quality seeds for further optimization}
We empirically study the capacity of \ourprog{} to improve the generation of seed sequences for subsequent local optimizations, showing an important application in biologically relevant scenario.

@@

\begin{figure}
  \hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering\includegraphics[width=0.9\textwidth]{Figs/td-widths}
  \vspace*{-24pt}
  \caption{Treewidths for multi-target RNA design instances of
    different complexity. Distributions of treewidths shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-widths}
\end{figure}

\begin{figure}
  \hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering
  \includegraphics[width=0.9\textwidth]{Figs/td-times}
  \vspace*{-24pt}
  \caption{Computation time of tree decompositions for
    multi-target RNA design instances of different complexity.
    Distributions of times (in ms/instance) shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-times}
\end{figure}

\begin{figure}[t!]
  %% the figure shows benchmark_inputs/RNAfold/3str/f3.100.0.inp
  \centering
  \verbatiminput{Figs/f3_100_0.txt}
  \includegraphicstop[width=0.6\textwidth]{Figs/td-example-basepair}%
  \includegraphicstop[width=0.35\textwidth]{Figs/td-example-stacking}
  \caption{Example instance from RF3 (top) with its tree decompositions in the base pair (left) and stacking model (right). The respective treewidths are 2 and 12.}
  \label{fig:td-example}
\end{figure}

@@

\begin{figure}
  \centering
  A)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/basepaircor}
  B)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/stackingcor}
  \caption{Validation of the trained parameters for the A) base pair
    model B) stacking model for predicting energies of the
    independently sampled sequences in the test set. We show
    correlation plots against the Turner energies reported by the
    ViennaRNA package.}
  \label{fig:training-cor}
\end{figure}

>>>>

\begin{figure*}[t]
%{\centering\includegraphics[width=.8\textwidth]{Figs/Workflow}\\}
\caption{General outline of \ourprog{} for base pair-based energy models. From a set of target secondary structures (i), base-pairs are merged (ii) into a compatibility graph (iii) and transformed into a tree decomposition (iv). The tree is then used to compute the partition function, followed by a Boltzmann sampling of valid sequences (v). An adaptive scheme learns weights to achieve targeted energies and $\Gb\Cb$-content, leading to the production of suitable designs (vi).}
\label{fig:workflow}
\end{figure*}

@@

To illustrate our above strategy, we sampled $n=1\,000$ sequences
targeting ${\sf GC}_{target}=0.5$ and different Turner energies for
the three structure targets of an example
instance. Figure~\ref{fig:energydist}A illustrates how the Turner
energy distributions of the shown target structures can be accurately
shifted to prescribed target energies; for comparison, we plot the
respective energy distribution of uniformly and Boltzmann sampled
sequences. Figure~\ref{fig:energydist}B summarizes the targeting accuracy for the single structure energies over a larger set of instances
from the \texttt{Modena} benchmark. We emphasize that only the simple stacking energies
are directly targeted (at $\pm$10\% tolerance, expect for a few hard
instances). Due to our linear adjustment, we still achieve well
controlled energies in the ultimately relevant Turner energy model.
\begin{figure*}[t]
  \begin{center}
   % {\sf \bfseries A}\includegraphicstop[width=0.57\textwidth]{Figs/energy_distribution}
    %{\sf \bfseries B}\includegraphicstop[width=0.32\textwidth]{Figs/offset}
  \end{center}
  \caption{%
  	Targeting specific energies using multi-dimensional Boltzmann sampling. (A) Turner energy distributions for three target structures (right) while targeting $(-40,-40,-20)$ and $(-20,-20,-20)$ free energy vectors. For comparison, we show the distribution of uniform and Boltzmann samples; respectively associated with homogenous weights $1$ and $e^\beta$. (B) Accuracy of targeting over all target structures of the \Software{Modena} benchmark instances \texttt{RNAtabupath}, \texttt{3str} and \texttt{4str}. Shown is the relative deviation of targeted and achieved energies in the simple stacking model and the Turner model.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
}
  \label{fig:energydist}
\end{figure*}
\subsection{Generating high-quality seeds for further optimization}
We empirically study the capacity of \ourprog{} to improve the generation of seed sequences for subsequent local optimizations, showing an important application in biologically relevant scenario.

@@

\begin{figure}
  \hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering%\includegraphics[width=0.9\textwidth]{Figs/td-widths}
  \vspace*{-24pt}
  \caption{Treewidths for multi-target RNA design instances of
    different complexity. Distributions of treewidths shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-widths}
\end{figure}

\begin{figure}
  \hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering
  %\includegraphics[width=0.9\textwidth]{Figs/td-times}
  \vspace*{-24pt}
  \caption{Computation time of tree decompositions for
    multi-target RNA design instances of different complexity.
    Distributions of times (in ms/instance) shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-times}
\end{figure}

\begin{figure}[t!]
  %% the figure shows benchmark_inputs/RNAfold/3str/f3.100.0.inp
  \centering
  \verbatiminput{Figs/f3_100_0.txt}
  %\includegraphicstop[width=0.6\textwidth]{Figs/td-example-basepair}%
  %\includegraphicstop[width=0.35\textwidth]{Figs/td-example-stacking}
  \caption{Example instance from RF3 (top) with its tree decompositions in the base pair (left) and stacking model (right). The respective treewidths are 2 and 12.}
  \label{fig:td-example}
\end{figure}

@@

\begin{figure}
  \centering
  A)%\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/basepaircor}
  B)%\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/stackingcor}
  \caption{Validation of the trained parameters for the A) base pair
    model B) stacking model for predicting energies of the
    independently sampled sequences in the test set. We show
    correlation plots against the Turner energies reported by the
    ViennaRNA package.}
  \label{fig:training-cor}
\end{figure}
